Statistical Theory\\ Chess Dataset Analysis

### 2024-09-29 Article/Article.tex
Updated Article

After these trasformations, we are left with 5,827 games which we will now begin to analyze.

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{ratingqq.png}
    \caption{Gamma-Poisson distribution fitted to the 'Turns' data}
    \label{fig:rating_qq}
\end{figure}

\subsubsection{Kolmogorov-Smirnov (KS) Test for Normality}
We applied the Kolmogorov-Smirnov (KS) test to evaluate whether the player rating distribution follows a normal distribution. The test returned a p-value of \(8.9649 \times 10^{-13}\), which is significantly lower than the threshold of \(0.03\) for a confidence level of 97\%, thus rejecting the null hypothesis that the rating distribution is normally distributed. However, when the test was performed on a random 10\% sample of player ratings, the p-value increased to 0.49, meaning we could not reject the null hypothesis at that sample size. This result indicates that while the entire population does not fit a normal distribution, smaller samples may approximate normality \cite{kssample}.

@@

\subsubsection{Mann-Whitney U Test for Game Length}
Since we could not confirm the normality of the rating distribution, we applied the non-parametric Mann-Whitney U test to compare the game lengths between high-rated and low-rated players. Our null hypothesis was that high-rated players do not play longer games than low-rated players on average. The test returned a p-value of \(2.7 \times 10^{-73}\), which allows us to reject the null hypothesis. Therefore, we conclude that player rating significantly influences game length, with higher-rated players tending to play longer games.

>>>>

After these trasformations, we are left with 16,155 games which we will now begin to analyze.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\linewidth]{ratingqq.png}
    \caption{Quantile - Quantile plot of player rating in relation to fitted normal distribution}
    \label{fig:rating_qq}
\end{figure}

\subsubsection*{Kolmogorov-Smirnov (KS) Test for Normality}
We applied the Kolmogorov-Smirnov (KS) test to evaluate whether the player rating distribution follows a normal distribution. The test returned a p-value of \(8.9649 \times 10^{-13}\), which is significantly lower than the threshold of \(0.03\) for a confidence level of 97\%, thus rejecting the null hypothesis that the rating distribution is normally distributed. However, when the test was performed on a random 10\% sample of player ratings, the p-value increased to 0.49, meaning we could not reject the null hypothesis at that sample size. This result indicates that while the entire population does not fit a normal distribution, smaller samples may approximate normality \cite{kssample}.

@@

@@

\subsubsection*{Mann-Whitney U Test for Game Length}
Since we could not confirm the normality of the rating distribution, we applied the non-parametric Mann-Whitney U test to compare the game lengths between high-rated and low-rated players. Our null hypothesis was that high-rated players do not play longer games than low-rated players on average. The test returned a p-value of \(2.7 \times 10^{-73}\), which allows us to reject the null hypothesis. Therefore, we conclude that player rating significantly influences game length, with higher-rated players tending to play longer games.
